```{r, include = FALSE, warning = FALSE}
# all the figures will be 6.5 x 4 inches and centered in the text.
knitr::opts_chunk$set(fig.width=6.5, 
                      fig.height=4, 
                      fig.align="center"
                      # fig.path="../figs/", 
                      )
## Set path to working directory
# knitr::opts_knit$set(root.dir = '../') 
# load packages
library(cdx2cea)
l_params_all <- load_all_params()
```

# Define model inputs {#inputs}

As described in the main manuscript, we declare all model input variables and set their values in this first component. The `R` script running the analysis of this component is the *01_model_inputs.R* file in the `analysis` directory. 

The input to inform the values is divided into three categories: external, estimated, and calibrated. External data inform the majority of the decision model parameters. Four parameter values need to be estimated using model calibration. 

In this component, we start with the general setup of the model, specifying the time horizon, name and number of health states, the proportion of the cohort in each of the different health states at the start of the simulation, and discount rates. The next step is to specify the external parameters. The initial model parameter values and `R` variable names are presented in Table \@ref(tab:parameters). 

Table: (\#tab:parameters) Description of the initial parameters with their `R` name and value of the decision model.

|           **Parameter**            |  **R name** |   **Value**   |
|:-----------------------------------|:------------|:--------------|
| Time horizon ($n_t$)               | `n_t`       | `r l_params_all$n_t` years |
| Names of health states ($n$)       | `v_n`       | H, S1, S2, D  |
| Annual discount rate (costs/QALYs) | `d_c`/`d_e` |  3%           |
| Annual transition probabilities    |             |               |
| - Disease onset (H to S1)          | `p_HS1`     |  `r l_params_all$p_HS1`         |
| - Recovery (S1 to H)               | `p_S1H`     |  `r l_params_all$p_S1H`          |
| - Disease progression (S1 to S2)  in the time-homogenous model   | `p_S1S2`    |  `r l_params_all$p_S1S2`        |
| Annual mortality                   |             |               |
| - All-cause mortality (H to D)     | `p_HD`      |  age-specific |
| - Hazard rate ratio of death in S1 vs H | `hr_S1`     |  `r l_params_all$hr_S1`            |
| - Hazard rate ratio of death in S2 vs H | `hr_S2`     |  `r l_params_all$hr_S1`           |
| Annual costs                       |             |               |
| - Healthy individuals              | `c_H`       |  $2,000       |
| - Sick individuals in S1           | `c_S1`      |  $4,000       |
| - Sick individuals in S2           | `c_S2`      |  $15,000      |
| - Dead individuals                 | `c_D`       |  $0           |
| - Additional costs of sick individuals treated in S1 or S2           | `c_Trt`     |  $12,000      |
| Utility weights                    |             |               |
| - Healthy individuals              | `u_H`       |  1.00         |
| - Sick individuals in S1           | `u_S1`      |  0.75         |
| - Sick individuals in S2           | `u_S2`      |  0.50         |
| - Dead individuals                 | `u_D`       |  0.00         |
| Intervention effect                |             |               |
| - Utility for treated individuals in S1 | `u_Trt` |  0.95        |

Age-specific background mortality for healthy individuals is represented by the US population in 2015 and obtained from the [Human Mortality database](https://www.mortality.org). This information is stored in the *01_all_cause_mortality.csv* file in the *data-raw* directory. Based on this .csv file a vector with mortality rates by age is created using the `load_mort_data` function in the *01_model_inputs_functions.R* script. This function gives us the flexibility to easily import data from other countries or years.

```{r}
print.function(load_mort_data) # print the function
``` 

Another function in the *01_model_inputs_functions.R* script, is the `load_all_parms` function. This function, which uses the `load_mort_data` function, loads all parameters for the decision model from multiple sources and creates a list containing all parameters and values.
```{r}
print.function(load_all_params)  # print the function
``` 

The `load_all_params` function is informed by the arguments `file.init` and `file.mort`. The `file.init` argument is a string with the file's location and name with the initial set of parameters. Our case-study's initial parameter values are stored in the *01_init_params.csv* file located in the *data-raw* directory. The `load_all_params` function read this .csv file into the function environment as a `data.frame` called, `df_params_init`.

The `file.mort` argument is a string with the file's location and name with mortality data. As described before, in our case-study, this is the *01_all_cause_mortality.csv* file. Within the `load_all_parms` function, the `load_mort_data` function is used to create a vector with mortality rates from the .csv data. 

After loading all the information, the `load_all_params` generates a list called `l_params_all,` including all the model parameters, including the general setup parameters and the vector of mortality rates. The function also stores the `data.frame` `df_params_init` with the initial set of parameters in the list. This is all executed in the *01_model_inputs.R* script by running the code below.
```{r, eval = FALSE, echo = TRUE}
l_params_all <- load_all_params()
```
We do not estimate parameters using individual-level data for the decision model, but rather, we estimate four parameters via model calibration. In this stage of the framework, we set these parameters to valid “dummy” values compatible with the next phase of the analysis, model implementation, but are ultimately just placeholder values until we conduct the calibration phase. This means that the best-fitted calibrated values will replace these values after performing the calibration in component 3.

Using a function to create a list of base-case parameters to have all model parameters in a single object is very useful because it will have to be updated for the calibration and the different sensitivity analyses in components 3 and 5 of the framework, respectively. Below, we guide you through the components of the function. 