
```{r, include = FALSE, warning = FALSE}
library(reshape2)
library(ggplot2)
library(cdx2cea)
### Load parameters
l_params_all <- load_all_params()
```

# Decision model {#simulation}

In this second component, we build the backbone of the decision analysis: implementing the model. This component is performed by the *02_decision_model.R* script. This file itself is not very large. It simply loads some packages, sources the input from component 01, and runs the `decision_model` function used to capture the decision model's dynamic process and stores the output. The output of the model is the traditional cohort trace. The trace describes how the cohort is distributed among the different health states over time and is plotted at the end of this script.

The function `decision_model` is defined in the *02_decision_model_functions.R* file in the `R` folder. As described in Alarid-Escudero et al. (2019)[@Alarid-Escudero2019e], constructing a model as a function facilitates subsequent stages of the model development and analysis. This process will call the same model function but pass different parameter values and/or calculate different outcomes based on the model outputs. In the next part, we will describe the code within the function.

```{r}
print.function(decision_model) # print the code of the function
```

The `decision_model` function is informed by the argument `l_params_all`. Via this argument, we give the function a list with all parameters of the decision model. These parameters are stored in the list `l_params_all`, which we passed into the function shown below.

```{r, eval = TRUE, echo = TRUE}
l_out_stm <- decision_model(l_params_all = l_params_all) # run the function
```

This function itself has all the mathematical equations of the decision models coded inside. It starts by calculating the age-specific transition probabilities from all non-dead states based on the vector of age-specific monthly mortality rates `v_r_mort_by_age_month`. These parameters will become vectors of length `n_cycles`, describing the probability to die for all ages from all non-dead states. 

The next part of the function creates an array that stores age-specific transition probability matrices in each third dimension. The transition probability matrix is a core component of a state-transition cohort model [@Iskandar2018]. This matrix contains the probabilities of transitioning from the current health state, indicated by the rows, to the other health states, specified by the columns. Since we have age-specific transition probabilities, the transition probability matrix is different each cycle. These probabilities are only depending on the cohort's age and not on other events; therefore, we can generate all matrices at the start of the model. This results in `n_cycles` different age-specific matrices that are stored in an array, called `a_P`, of dimensions `n_states` $\times$ `n_states` $\times$ `n_cycles`. After initializing the array, it is filled with the transition probability stored in the list. When running the model, we can index the correct transition probability matrix corresponding with the cohort's current age. We then added some sanity checks to ensure that the transition matrices and the transition probabilities are valid. The first three and last cycles of the transition probability matrices stored in the array `a_P` are shown below.

```{r, echo = TRUE, eval = TRUE}
l_out_stm$a_P[, , 1:3] # show the first three time-points of a_P
l_out_stm$a_P[, , l_params_all$n_cycles] # show it for the last cycle
```
By comparing these probability matrices, we observe an increase in the probabilities of transitioning to death from all health states. 

After the array is filled, the cohort trace matrix, `m_M`, of dimensions `n_cycles` $\times$ `n_states` is initialized. This matrix will store the state occupation at each point in time. The initial state vector `v_s_init` informs the first row of the matrix. We iteratively multiply the cohort trace with the age-specific transition probability matrix corresponding to the specific cycle obtained by indexing the array `a_P` appropriately for the remaining points in time. All the decision model's outputs and relevant elements are stored in a list called `l_out_stm`. This list contains the array of the transition probability matrix for all cycles `t` and the cohort trace `m_M`. 

```{r, eval = TRUE, echo = TRUE}
head(l_out_stm$m_M)    # show the top part of the cohort trace
tail(l_out_stm$m_M)    # show the bottom part of the cohort trace
```

Using the code below, we can graphically show the model dynamics by plotting the cohort trace. Figure \@ref(fig:Cohort-Trace) shows the cohort's distribution among the different health states at each time point.

```{r Cohort-Trace, echo = FALSE, fig.cap = 'Cohort trace of the cohort model', message = FALSE, warning = FALSE, fig.pos = "H"} 
ggplot(melt(l_out_stm$m_M), aes(x = Var1, y = value, color = Var2, linetype = Var2)) +
  geom_line(size = 1.3) +
  scale_color_discrete(l = 50, name = "Health state") +
  scale_linetype_discrete(name = "Health state") +
  scale_y_continuous(limits = c(0, 1)) +
  xlab("Cycle") +
  ylab("Proportion of the cohort") +
  theme_bw(base_size = 16) +
  theme()
```

